---
title: "Guided Exercise 12 - Databases"
output: 
  html_document: default
---

---
title: "Guided Exercise 12 - Databases"
output: 
  tufte::tufte_html:
    css: ../ex.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(tidyverse)
theme_set(theme_bw())
options(scipen = 5) # encourage metrics to print in fixed-point notation
options(dplyr.summarise.inform = FALSE) # silence a warning message
```

```{r}
library(tidyverse)
library(mdsr)
library(RMySQL)
db <- DBI::dbConnect(RMariaDB::MariaDB(), dbname = "airlines", host = "mdsr.cdc7tgkkqd0n.us-east-1.rds.amazonaws.com", 
    user = "mdsr_public", password = "ImhsmflMDSwR")
```

And we'll get references to two specific tables we'll be working with.

```{r}
flights <- tbl(db, "flights")
carriers <- tbl(db, "carriers")
```

```{r}
flights %>% 
  filter(year == 2016) %>% 
  show_query()
```

```{r}
nrow(flights)
```


##Exercise 1
WHERE corresponds to filter

```{r}
flights %>% 
  filter(year == 2016, month == 11, distance > 1000) %>% 
  summarize(
    n = n(),
    mean_dep_delay = mean(dep_delay, na.rm = TRUE)
  ) %>% show_query
```


##Exercise 2
It uses COUNT, AVG when selecting the columns.

##Exercise 3
```{r}

flights %>% 
  summarize(minyear = min(year), maxyear = max(year) )
```

##Exercise 4
```{r}

flights %>% 
  filter(dest == 'JFK', year == 2016, month == 11) %>% 
  summarize(n = n())
```


##Exercise 5
```{r}

ex5 <- flights %>% 
  filter(dest == 'JFK', year == 2016, month == 11) %>% 
  group_by(carrier) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
ex5
```
##Exercise 6
```{r}
ex5 %>% show_query
```
SQL expresses arrange by using ORDER BY

##Exercise 7
```{r}
sum_flights <- flights %>% 
  filter(dest == 'JFK', year == 2016, month == 11) %>% 
  group_by(carrier) %>% 
  summarize(n = n(), frac_ontime = sum(arr_delay<= 15)/n()) %>% 
  arrange(desc(n))
sum_flights
```

##Exercise 8
```{r}
sum_flights %>% 
  inner_join(carriers, sum_flights, by = 'carrier') %>% 
  arrange(desc(n))

```
```{r}
carriers
```

