---
title: "Homework 3: Bike Sharing Data Wrangling"
author: "Daniel Kwik"
output: prettydoc::html_pretty
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
# Import some functions from lubridate using less confusing names
get_year <- lubridate::year; get_month <- lubridate::month; get_day <- lubridate::day; get_weekday <- lubridate::wday

# Disable a warning message.
options(dplyr.summarise.inform = FALSE)
```

## Load Data

```{r load-data}
# Read the file, specifying data types for each column.
rides_2011 <- read_csv("data/rides_2011.csv.xz", col_types = cols(
  start_time = col_datetime(),
  duration = col_double(),
  rider_type = col_factor()
))
```

## Exercise 1

Rides contains 3 columns, start_time, duration, and rider_type. Start_time describes the date and time of the ride, duration contains numbers that is related to how long the ride was, and the rider_type column describes the type of rider as either 'Casual' or 'Member'.

## Exercise 2

```{r}
rides_2011 %>% 
  arrange(desc(duration)) %>% 
  mutate(duration_hours = (duration/60/60))
```

The duration column is probably in seconds. Arranging by descending duration and dividing by 60 and then by 60 again gives us a max of 23.98 units in duration. This is likely hours because the system does not allow durations longer than 24 hours. Since we divide by 60 twice to get hours, duration must originally be in seconds.

## Exercise 3

```{r}
rides_2011 %>% 
  group_by(rider_type) %>% 
  summarise(rides = n()) %>% 
  knitr::kable()
```

## Exercise 4

```{r}
daily_rides <- rides_2011 %>%
  mutate(year = get_year(start_time),
         month = get_month(start_time),
         day = get_day(start_time)) %>% 
  group_by(year, month, day) %>% 
  summarize( rides = n())

daily_rides
  
```

## Exercise 5

daily_rides has `r nrow(daily_rides)` rows, and it should have 365 rows.

## Exercise 6

```{r}
daily_rides_by_type <- rides_2011 %>%
  mutate(year = get_year(start_time),
         month = get_month(start_time),
         day = get_day(start_time)) %>% 
  group_by(year, month, day, rider_type) %>% 
  summarize( rides = n())
daily_rides_by_type
  
```

## Exercise 7

```{r}
rides_over_week <- rides_2011 %>%
  mutate(year = get_year(start_time),
         month = get_month(start_time),
         day = get_day(start_time),
         weekday = get_weekday(start_time, label = TRUE)) %>% 
  group_by(year, month, day, rider_type, weekday) %>% 
  summarize( rides = n()) %>% 
  filter(rider_type != "Unknown")
rides_over_week

ggplot(rides_over_week, aes(x = weekday, y = rides, fill = rider_type)) +
  geom_boxplot() +
  labs(x="Day of Week", 
       y = "Rides (thousands)", 
       fill = "Type of Rider", 
       title = "Distribution of rides by day of week")
```

## Exercise 8

```{r}
rides_over_week %>% 
  group_by(weekday, rider_type) %>% 
  summarize(median_rides = median(rides)) %>% 
  pivot_wider(names_from = rider_type, values_from = median_rides) %>% 
  knitr::kable()
```
