---
title: "Exercise 05 - Visualizing Spatial Data"
author: "Daniel Kwik"
output:
  prettydoc::html_pretty
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
dennys <- read_rds("data/dennys.rds")
laquinta <- read_rds("data/laquinta.rds")
states <- read_csv("data/states.csv", col_types = cols(
  name = col_character(),
  abbreviation = col_character(),
  area = col_double()
))
```

### Exercise 1

```{r}
glimpse(dennys)

```

The dimensions are `r nrow(dennys)` rows and `r ncol(dennys)` columns.


### Exercise 2
```{r}
glimpse(laquinta)
```
There are 999 rows and 6 columns. Each row represents a single address, including the variables city, state, zip code, and coordinates.


### Exercise 3
```{r}
dennys %>% 
  filter(!(state%in% states$abbreviation))
```
There are no Denny's locations outside the US.

### Exercise 4

```{r}
dennys <- dennys %>% 
mutate( country = "United States")
```


### Exercise 5

```{r}
laquinta %>% 
  filter(!(state %in% states$abbreviation))
```

### Exercise 6
```{r}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL") ~ "Mexico",
    state == "FM" ~ "Honduras"))

laquinta <- laquinta %>%
  filter(country == "United States")

```


### Exercise 7
```{r}
dennys %>% 
  group_by(state) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
dennys

```
```{r}

laquinta %>% 
  group_by(state) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count))
laquinta
```


### Exercise 8
```{r}
dennys %>% 
  group_by(state) %>% 
  summarize(count=n()) %>% 
  arrange(desc(count)) %>% 
  left_join(states, by = c("state" = "abbreviation")) %>% 
  mutate(count_by_area = count / area)
dennys
```

### Exercise 9

### Exercise 10

### Exercise 11
